Loblaws Seasonal Discount Patterns: Peak Sales Frequency in Fall and Deeper Discounts in Early Spring

### 2024-11-15 paper/paper.qmd
committing formatting changes to paper.qmd

## Discount Distribution by Season

@fig-seasonal-discount-distrib highlights the spread and density of discount percentages for each season. The violin plot shows that fall has the widest range of discount percentages, with some discounts reaching high values, indicating that the fall season sees a more aggressive discounting strategy. In contrast, winter displays a narrower distribution, suggesting more conservative or consistent discounting during this period. The density of moderate discounts in spring and summer implies steady discount practices, with fewer extreme discounts than seen in fall. Overall, this distribution analysis suggests that Loblaws adjusts its discount depth based on seasonal factors, with potentially deeper or more varied discounts offered in the fall.
```{r}
#| echo: false
#| eval: true
#| label: fig-seasonal-discount-distrib
#| fig-cap: "Distribution of discount percentages across seasons at Loblaws, indicating higher variability and deeper discounts in fall."
#| warning: false

ggplot(price_analysis_data %>% filter(on_sale == TRUE), aes(x = season, y = discount_percentage)) +
  geom_violin(fill = "lightblue", color = "black") +
  labs(
    x = "Season",
    y = "Discount Percentage"
  ) +
  theme_minimal()
```

@@

@fig-average-monthly-discount provides insight into how average discount rates vary across months. The chart reveals that the highest average discounts occur in early spring (February and March) and gradually decrease throughout the year, stabilizing at a lower average discount from July onward. This trend may reflect early-year inventory clearing, where Loblaws offers more substantial discounts to move products after the holiday season. The stabilization of discounts in the latter half of the year, despite an increase in sale frequency, suggests that while more items are on sale, the depth of individual discounts is reduced. This balance between volume and discount depth likely reflects an optimized strategy for attracting steady consumer interest without significant margin reductions.
```{r}
#| echo: false
#| eval: true
#| label: fig-average-monthly-discount
#| fig-cap: "Average discount percentage per month, showing the highest discounts in early spring, with a gradual decline through the year."
#| warning: false

@@

# Plot the data
ggplot(avg_discount_by_month, aes(x = month, y = avg_discount)) +
     geom_point(color = "blue", size = 3) +
     labs(
         x = "Month",
         y = "Average Discount Percentage"
     ) +
     theme_minimal()
```

@@

ggplot(price_analysis_data %>% filter(on_sale == TRUE), aes(x = discount_percentage, fill = season)) +
  geom_density(alpha = 0.5) +
  labs(
    x = "Discount Percentage",
    y = "Density"
  ) +
  theme_minimal()
```
Collectively, the visualizations provide insights into Loblaws' seasonal and monthly discount strategies. The data suggests that Loblaws increases the frequency of sales in the latter half of the year while moderating the average discount depth, likely balancing consumer demand with profit margins. The fall season appears to be a critical period for aggressive discounting, likely tied to seasonal shopping behaviors. By analyzing both the frequency and depth of discounts, this study highlights the complexity of retail discount strategies and provides a basis for further investigation into how these strategies impact consumer behavior and sales outcomes.

>>>>

## Discount Distribution by Season
```{r}
#| echo: false
#| eval: true
#| label: fig-seasonal-discount-distrib
#| fig-cap: "Distribution of discount percentages across seasons at Loblaws, indicating higher variability and deeper discounts in fall."
#| warning: false

ggplot(price_analysis_data %>% filter(on_sale == TRUE), aes(x = season, y = discount_percentage)) +
  geom_violin(fill = "lightblue", color = "black") +
  labs(
    x = "Season",
    y = "Discount Percentage"
  ) +
  theme_minimal()
```
@fig-seasonal-discount-distrib highlights the spread and density of discount percentages for each season. The violin plot shows that fall has the widest range of discount percentages, with some discounts reaching high values, indicating that the fall season sees a more aggressive discounting strategy. In contrast, winter displays a narrower distribution, suggesting more conservative or consistent discounting during this period. The density of moderate discounts in spring and summer implies steady discount practices, with fewer extreme discounts than seen in fall. Overall, this distribution analysis suggests that Loblaws adjusts its discount depth based on seasonal factors, with potentially deeper or more varied discounts offered in the fall.

@@

```{r}
#| echo: false
#| eval: true
#| label: fig-average-monthly-discount
#| fig-cap: "Average discount percentage per month, showing the highest discounts in early spring, with a gradual decline through the year."
#| warning: false

@@

# Plot the data
ggplot(avg_discount_by_month, aes(x = month, y = avg_discount)) +
     geom_point(color = "blue", size = 3) +
     labs(
         x = "Month",
         y = "Average Discount Percentage"
     ) +
     theme_minimal()
```
@fig-average-monthly-discount provides insight into how average discount rates vary across months. The chart reveals that the highest average discounts occur in early spring (February and March) and gradually decrease throughout the year, stabilizing at a lower average discount from July onward. This trend may reflect early-year inventory clearing, where Loblaws offers more substantial discounts to move products after the holiday season. The stabilization of discounts in the latter half of the year, despite an increase in sale frequency, suggests that while more items are on sale, the depth of individual discounts is reduced. This balance between volume and discount depth likely reflects an optimized strategy for attracting steady consumer interest without significant margin reductions.

@@

ggplot(price_analysis_data %>% filter(on_sale == TRUE), aes(x = discount_percentage, fill = season)) +
  geom_density(alpha = 0.5) +
  labs(
    x = "Discount Percentage",
    y = "Density"
  ) +
  theme_minimal()
```

Collectively, the visualizations provide insights into Loblaws' seasonal and monthly discount strategies. The data suggests that Loblaws increases the frequency of sales in the latter half of the year while moderating the average discount depth, likely balancing consumer demand with profit margins. The fall season appears to be a critical period for aggressive discounting, likely tied to seasonal shopping behaviors. By analyzing both the frequency and depth of discounts, this study highlights the complexity of retail discount strategies and provides a basis for further investigation into how these strategies impact consumer behavior and sales outcomes.
